Dynamics of Laterally Propagating Flames in X-ray Bursts

### 2019-10-14 xrb/paper.tex
minor changes

Figure~\ref{fig:time_series} shows the time evolution of the 10/10
simulation, focusing on the mean-molecular weight,
\begin{equation}
\bar{A} = \left ( \sum_k \frac{X_k}{A_k} \right )^{-1}
\end{equation}
In each frame the buffer of \isot{Ni}{56} that serves as the
underlying neutron star is seen spanning the bottom of the domain.
Above that, the composition begins as pure \isot{He}{4}, but as the
simulation progresses, the flame processes this to heavier nuclei,
increasing $\bar{A}$.  By about $10~\mbox{ms}$, we see the flame front
is reasonable well-defined.  We see that the bottom of the burning
front is listed off of the base of the atmosphere, greatly increasing
the surface area of the burning compared with a perfectly vertical
flame front.  By $20~\mbox{ms}$ the flame has moved out substantially
and we are beginning to see a gradient in the composition of the ash,
with the heavier nuclei furthest behind the flame.  The boosting of
the burning likely artificially increases this effect.

@@

Figure~\ref{fig:10_10_overview} shows the temperature, energy
generation rate, and $w$ component of the velocity (the out-of-plane
velocity induced by the Coriolis force).  This latter field
illustrates the hurricane effect setup by the laterally spreading
burning front.  In the energy generation rate plot, we see that the
burning is mostly concentrated toward the bottom of the layer, as
expected since the density is greatest there.  We see that the peak of
the burning has moved off of the symmetry axis, demonstrating that the
burning front is propagating to the right.  In the temperature plot,
we see the effect of our refinement criteria focusing only on the
atmosphere where we are most dense, with an artificial drop in the
temperature at the refinement boundary likely due to the thermal
diffusion.  As we'll see when the lower resolution case, we do not
believe that this affects the results.

A final feature worth noting is the ash that seems to move along the
surface at a higher velocity than the flame, via surface gravity
waves.  The sponging that we perform is likely damping this some, and
the method by which we initialize the flame may induce a larger
transient than in nature, nevertheless, this surface ash is intriguing
because it affects the composition of the photosphere ahead of the
burning front, potentially changing our interpretation of
observations.  This is something that will be explored more fully in
the future.

@@

Our default resolution puts $\sim 80$ zones vertically in the
``cool'' \MarginPar{double check} model atmosphere height.  To
understand the effects of resolution, we also perform a run with one
fewer level, giving at $20~\mbox{cm}$ resolution overall (this is our
10/10-lores run).  Figure~\ref{fig:10_10_lowres} shows the fields at
20~ms.  The structure is largely the same as the 10/10 run, with
largely the same flame shape and position, and the same structure in
the energy generation rate.  Since we do not have a jump in refinement
right below the atmosphere, we don't see the strong diffusive effect
there, but we do see some cooling in the \isot{Ni}{56} region.  This
is likely a resolution effect, again due to the strong gradient in
temperature at the base of the atmosphere.

@@

We show some phase plots in Figure~\ref{fig:phase_plots}. The phase plot of the
energy generation rate against the as a function of
the $x$- and $y$-velocities shows that energy is preferentially generated in
regions with negative $x$-velocity. This most likely corresponds to material
being entrained along the underside of the flame front. There are loops of
points of similar $\dot{e}_\mathrm{nuc}$ in the outer edges of the plot.
These are likely to correspond to the vortices that appear within the flame,
and show the fluid moving in a circular motion in $u-v$ phase space.

The phase plot of the energy generation rate as a function of the $x$-velocity
and the density demonstrates that the energy generation rate peaks at
$\rho \sim 3 \times 10^5~\mathrm{g/cm}^3$, at the base of the flame.
\MarginPar{I'm guessing? It's at least towards the base}

\begin{figure}[t]
\centering
\plottwo{plots/results/vx_vy_129139}{plots/results/u_rho_129139}
\caption{\label{fig:phase_plots} Phase plots at time $t = 0.02~\mathrm{s}$.
\emph{Left}: Phase plot showing the energy generation rate as a function of
the $x$- and $y$-velocities. \emph{Right}: phase plot showing the energy
generation rate as a function of the $x$-velocity and the density.}
\end{figure}

>>>>

Figure~\ref{fig:time_series} shows the time evolution of the 10/10
simulation, focusing on the mean-molecular weight,
\begin{equation}
\bar{A} = \left ( \sum_k \frac{X_k}{A_k} \right )^{-1}
\end{equation}
In each frame the buffer of \isot{Ni}{56} that serves as the
underlying neutron star is seen spanning the bottom of the domain.
Above that, the composition begins as pure \isot{He}{4}, but as the
simulation progresses, the flame processes this to heavier nuclei,
increasing $\bar{A}$.  By about $10~\mbox{ms}$, we see the flame front
is reasonable well-defined.  We see that the bottom of the burning
front is lifted off of the base of the atmosphere, greatly increasing
the surface area of the burning compared with a perfectly vertical
flame front.  By $20~\mbox{ms}$ the flame has moved out substantially
and we are beginning to see a gradient in the composition of the ash,
with the heavier nuclei furthest behind the flame.  The boosting of
the burning likely artificially increases this effect. \MarginPar{Doesn't the increased flame speed of the 10/10 run vs. the 5/5 run show this?}

@@

Figure~\ref{fig:10_10_overview} shows the temperature, energy
generation rate, and $w$ component of the velocity (the out-of-plane
velocity induced by the Coriolis force).  This latter field
illustrates the hurricane effect set up by the laterally spreading
burning front.  In the energy generation rate plot, we see that the
burning is mostly concentrated toward the bottom of the layer, as
expected since the density is greatest there.  We see that the peak of
the burning has moved off of the symmetry axis, demonstrating that the
burning front is propagating to the right.  In the temperature plot,
we see the effect of our refinement criteria focusing only on the
atmosphere where we are most dense, with an artificial drop in the
temperature at the refinement boundary likely due to the thermal
diffusion. \MarginPar{Remind the reader what this was/where they should be looking in the plot to see this temperature drop} As we'll see when the lower resolution case, we do not
believe that this affects the results.

A final feature worth noting is the ash that seems to move along the
surface at a higher velocity than the flame, via surface gravity
waves.  The sponging that we perform is likely damping this to some extent, and
the method by which we initialize the flame may induce a larger
transient than in nature, nevertheless, this surface ash is intriguing
because it affects the composition of the photosphere ahead of the
burning front, potentially changing our interpretation of
observations.  This is something that will be explored more fully in
the future.

@@

Our default resolution puts $\sim 80$ zones vertically in the
``cool'' \MarginPar{double check} model atmosphere height.  To
understand the effects of resolution, we also performed a run with one
fewer level, giving a $20~\mbox{cm}$ resolution overall (this is our
10/10-lores run).  Figure~\ref{fig:10_10_lowres} shows the fields at
20~ms.  The structure is largely the same as the 10/10 run, with
largely the same flame shape and position, and the same structure in
the energy generation rate.  Since we do not have a jump in refinement
right below the atmosphere, we don't see the strong diffusive effect
there, but we do see some cooling in the \isot{Ni}{56} region.  This
is likely a resolution effect, again due to the strong gradient in
temperature at the base of the atmosphere.

@@

We show some phase plots in Figure~\ref{fig:phase_plots}. The phase plot of the
energy generation rate as a function of
the $r$- and $z$-velocities shows that energy is preferentially generated in
regions with negative $r$-velocity. This most likely corresponds to the burning along the underside of the flame front, where fuel has been entrained along the surface of the flame. This is further supported by the phase plot of the energy generation rate as a function of the $r$-velocity
and the density, which shows that the energy generation rate peaks at
$\rho \sim 3 \times 10^5~\mathrm{g/cm}^3$, at the base of the flame.

In the first plot, there are ``loops" of
points of similar $\dot{e}_\mathrm{nuc}$ in the outer edges of the plot.
These are likely to correspond to the vortices that appear within the flame,
and show the fluid moving in a circular motion in $u-v$ phase space.

\begin{figure}[t]
\centering
\plottwo{plots/results/vx_vy_129139}{plots/results/u_rho_129139}
\caption{\label{fig:phase_plots} Phase plots at time $t = 0.02~\mathrm{s}$.
\emph{Left}: Phase plot showing the energy generation rate as a function of
the $r$- and $z$-velocities. The black cross shows the location of $u = v = 0$. \emph{Right}: phase plot showing the energy
generation rate as a function of the $r$-velocity and the density.}
\end{figure}
